# Step 1: soil data preparation
This chapter builds on the previous one as it requires basic understanding of the data handling using **R**. From this point onwards, the steps base on each other and are needed to complete the mapping process. The instructions covered in this chapter provide step-by-step instructions on the following items:

1. Perform a quality check of the data
2. Estimate bulk density using PTF
<!-- * estimate organic carbon stocks (OCS) -->
3. Harmonize soil layers (using splines)
4. Plot and save the formatted soil data

## Load national data
As specified in the previous Chapter in regards to [pre-processing](preproc), at first the necessary steps in *RStudio* are taken: set the working directory and load the necessary R packages. Note that there are many ways to install packages besides the most common way using the function `install.packages()`. For instance, to install the `terra` package, one has to write `install.packages("terra")`. This installs the package from CRAN. However, there are a few exceptions where development versions of R packages are required. In these instances additional packages such as `devtools` or `remotes` are needed (see example in code below). These packages are then able to install packages from for instance GitHub repositories.

```{r, eval = F}
# 0 - User-defined variables ===================================================
#wd <- 'C:/Users/luottoi/Documents/GitHub/GSNmap-TM/Digital-Soil-Mapping'
#wd <- "C:/GIT/GSNmap-TM/Digital-Soil-Mapping"

# 1 - Set working directory and load necessary packages ========================
setwd(wd) # change the path accordingly

library(tidyverse) # for data management and reshaping
library(readxl) # for importing excel files
library(mapview) # for seeing the profiles in a map
library(sf) # to manage spatial data (shp vectors) 
library(aqp) # for soil profile data
#install.packages("devtools") 
#devtools::install_bitbucket("brendo1001/ithir/pkg") #install ithir package
library(ithir) # for horizon harmonization
library(data.table)

```

The next step is to load the national soil data into *R Studio*. For that, it is recommendable to have the data in either Microsoft Excel format (.xlsx) or as comma separated value table (.csv). In both cases, each row represents a sample (or horizon) and each column represents a variable. Then, the datasets can be loaded from the specified folder using the respective functions specified in the code below. It is noteworthy that in **R** datasets also need to be assigned to a user-defined variable in order to be saved in the "global environment".

In this example, the three different data tables are loaded into *RStudio*. The soil profile database of SISINTA (`hor`), the chemical (`chem`) and physical soil property tables (`phys`). After reading in the file, the package `tidyverse` comes into play. By using the `select()` and `unique()` functions, the user can select only the necessary columns from the table and ensure that no duplicates are included. At this point it may be necessary to rename certain columns, as shown for the Profile and Horizon ID columns in the code below.
Finally, every time new datasets are loaded into *RStudio*, it is recommendable to check the data. Using the `summary()` function, users can see the class of each variable (= column) and descriptive statistics (for numerical variables). Classes are 'character' (`chr`) for text, integer (`int`) for whole numbers, and numeric (`num`) for numeric variables. 


```{r, eval = T, message=F, warning=F}
# 2 - Import national data =====================================================
# Save your national soil dataset in the data folder /01-Data as a .csv file or 
# as a .xlsx file

## 2.1 - for .xlsx files -------------------------------------------------------
# Import horizon data 
# hor <- read_excel("01-Data/soil_data.xlsx", sheet = 2)
# # Import site-level data
# site <- read_excel("01-Data/soil_data.xlsx", sheet = 1)
# chem <- read_excel("01-Data/soil_data.xlsx", sheet = 2)
# phys <- read_excel("01-Data/soil_data.xlsx", sheet = 3)


## 2.2 - for .csv files --------------------------------------------------------
# Import horizon data 
hor <- read_csv(file = "Digital-Soil-Mapping/01-Data/soil_profile_data.csv")
chem <- read_csv(file = "Digital-Soil-Mapping/01-Data/soil_chem_data030.csv")
phys <- read_csv(file = "Digital-Soil-Mapping/01-Data/soil_phys_data030.csv")

site <- select(hor, id_prof, x, y) %>% unique()
hor <- select(hor, id_prof, id_hor, top:cec)

# change names of key columns
names(site)
names(site)[1] <- "ProfID"
names(hor)
names(hor)[1] <- "ProfID"
names(hor)[2] <- "HorID"
# scan the data
summary(site)
summary(hor)

```

The selection of useful columns is very important since it ensures that users keep a good overview and a clean environment. Using the `select()` function, it is also possible to rename the variables right away (see code below).

```{r select, eval = T}
# 3 - select useful columns ====================================================
## 3.1 - select columns --------------------------------------------------------
hor <- select(hor, ProfID, HorID, top, bottom, ph=ph_h2o, k, soc, bd, cec)

```


## Data quality check

Datasets need to be checked for their quality as especially manually entered data is prone to mistakes such as typos or duplicates. A thorough quality check ensures that:

* all profiles have reasonable coordinates (within the area of interest);
* there are no duplicated profiles; and
* the depth logic within a profile is not violated.

To check the first point, the dataframe needs to be converted into a spatial object using the `st_as_sf()` function of the `sf` package. It is necessary to indicate the columns that contains latitude and longitude, as well as a coordinate reference system (CRS). We recommend WGS84 which corresponds to an EPSG code of 4326. However, locally more appropriate CRS can be found on the following website: https://epsg.io/. The `mapview()` command (from `mapview` package) offers the possibility to visualize the profile locations in an interactive map. Finally, the `filter()` function can be used to remove rows that contain profiles with wrong locations.

```{r aqp, echo=FALSE, message=FALSE}
# 4 - Quality check ============================================================

## 4.1 - Check locations -------------------------------------------------------
# https://epsg.io/4326
site %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326) %>% # convert to spatial object
  mapview(zcol = "ProfID", cex = 3, lwd = 0.1) # visualise in an interactive map

# profile 2823 is wrongly located, so let's remove it
site <- filter(site, ProfID != 2823)

```

To visualize the profile locations, the soil data table was converted into a shapefile. Still, to check whether the database complies with the depth logic within each profile, it is necessary to convert the data table into a so-called soil profile collection that allows for very specific operations. These operations were bundled in the package `aqp` (AQP = Algorithms for Quantitative Pedology) [@beaudette2013].
With the first lines of code below, the dataset is converted into a soil profile collection and profiles and horizon tables are joined based on the site information.
Now the profile collection can be visualised for any soil property. In this case, only the first 20 profiles are selected for the cation exchange capacity (CEC).
Using the `checkHzDepthLogic()` function, users can assess that all profiles do not have gaps or overlaps of neighbouring horizons. 


```{r, eval = T}
## 4.2 - Convert data into a Soil Profile Collection ---------------------------
library(aqp)
depths(hor) <- ProfID ~ top + bottom
hor@site$ProfID <- as.numeric(hor@site$ProfID)
site(hor) <- left_join(site(hor), site)
profiles <- hor

profiles

## 4.3 - plot first 20 profiles using pH as color ------------------------------
plotSPC(x = profiles[1:20], name = "cec", color = "cec",
        name.style = "center-center")

## 4.4 - check data integrity --------------------------------------------------
# A valid profile is TRUE if all of the following criteria are false:
#    + depthLogic : boolean, errors related to depth logic
#    + sameDepth : boolean, errors related to same top/bottom depths
#    + missingDepth : boolean, NA in top / bottom depths
#    + overlapOrGap : boolean, gaps or overlap in adjacent horizons
aqp::checkHzDepthLogic(profiles)

```

If there are profiles that violate the depth logic rules (i.e. overlapping horizons), they can be selected and checked through the Profile ID. In the following step, only profiles with valid horizon logic are selected. Finally, the soil profile collection is re-converted to a dataframe. With this, the quality check is finished.

```{r, eval = T}
# visualize some of these profiles by the pid
subset(profiles, grepl(6566, ProfID, ignore.case = TRUE))
subset(profiles, grepl(6915, ProfID, ignore.case = TRUE))
subset(profiles, grepl(7726, ProfID, ignore.case = TRUE))


## 4.5 - keep only valid profiles ----------------------------------------------
clean_prof <- HzDepthLogicSubset(profiles)
metadata(clean_prof)$removed.profiles
# write_rds(clean_prof, "01-Data/soilProfileCollection.rds")

## 4.6 convert soilProfileCollection to a table --------------------------------
dat <- left_join(clean_prof@site, clean_prof@horizons)
dat <- select(dat, ProfID, HorID, x, y, top, bottom, ph:cec )

```


## Calculation of pedo-transfer functions

In the cases of single-layer samples, which is common in sampling for nutrient determination, a locally calibrated pedotransfer function (PTF) should be applied. PTF will be also required to harmonise the laboratory methods. Experts from GLOSOLAN will provide advice in this regard.

Therefore, a customised function is introduced to our working environment. Users can write their own functions in **R**. This is often necessary when existing functions need to be customised or very specific calculations need to be performed. Functions greatly increase the efficiency of our code. For further information, it is recommendable to consult online resources on the topic (e.g. https://hbctraining.github.io/Intro-to-R/lessons/03_introR-functions-and-arguments.html).

The function `estimateBD` below calculates various PTFs that estimate BD. Which equation is used is determined by the user that has to choose one of the methods and also specify the SOC value of the respective horizon. The SOC values is first converted to OM by using the conversion factor of 1.724 and then inserted in the respective PTF. The return() command tells **R** which value to output. 

```{r, eval = T, message = F}
# 5 - Estimate BD using pedotransfer functions =================================

# create the function with all PTF
estimateBD <- function(SOC=NULL, method=NULL){
  OM <- SOC * 1.724
  if(method=="Saini1996"){BD <- 1.62 - 0.06 * OM}
  if(method=="Drew1973"){BD <- 1 / (0.6268 + 0.0361 * OM)}
  if(method=="Jeffrey1979"){BD <- 1.482 - 0.6786 * (log(OM))}
  if(method=="Grigal1989"){BD <- 0.669 + 0.941 * exp(1)^(-0.06 * OM)}
  if(method=="Adams1973"){BD <- 100 / (OM /0.244 + (100 - OM)/2.65)}
  if(method=="Honeyset_Ratkowsky1989"){BD <- 1/(0.564 + 0.0556 * OM)}
  return(BD)
}
```

To apply the `estimateBD` function, first a test dataframe is created that includes the SOC values from the cleaned profile table as well as the respective existing BD measurements. The rows without values in one of the columns are excluded using the na.omit() function since we want to first evaluate the difference between estimated BDs and measured BDs. 
Now, the test dataframe is complemented by the estimated BDs derived from the PTFs for each method. To add new columns to an existing dataframe one has to write on the left-hand side of the arrow the name of the existing dataframe object (in this case BD_test), the dollar sign ($), and the name of the new column. Here, the names are given according to the used BD PTF.

```{r, eval = T, message = F}
## 5.1 - Select a pedotransfer function ----------------------------------------
# create a vector of BD values to test the best fitting pedotransfer function
BD_test <- tibble(SOC = clean_prof@horizons$soc,
                  BD_test = clean_prof@horizons$bd)
BD_test <-  na.omit(BD_test)

## 5.2 - Estimate BLD for a subset using the pedotransfer functions ------------
BD_test$Saini <- estimateBD(BD_test$SOC, method="Saini1996")
BD_test$Drew <- estimateBD(BD_test$SOC, method="Drew1973")
BD_test$Jeffrey <- estimateBD(BD_test$SOC, method="Jeffrey1979")
BD_test$Grigal <- estimateBD(BD_test$SOC, method="Grigal1989")
BD_test$Adams <- estimateBD(BD_test$SOC, method="Adams1973")
BD_test$Honeyset_Ratkowsky <- estimateBD(BD_test$SOC,
                                         method="Honeyset_Ratkowsky1989")

```

The calculated BDs can now be compared using the `summary()` function. However, a faster and more accessible approach is to plot the different bulk densities for comparison. In case you are not familiar with the `plot()` function and its respective commands, it is recommendable to check one of the many online learning resources such as https://intro2r.com/simple-base-r-plots.html. The plot shows us both measured and estimated BD values as differently coloured lines. 

```{r, eval = T}
## 5.3 Compare results ---------------------------------------------------------

# Observed values:
summary(BD_test$BD_test)

# Predicted values:
summary(BD_test$Saini)
summary(BD_test$Drew)
summary(BD_test$Jeffrey)
summary(BD_test$Grigal)
summary(BD_test$Adams)
summary(BD_test$Honeyset_Ratkowsky)

# Compare data distributions for observed and predicted BLD
plot(density(BD_test$BD_test),type="l",col="black", ylim=c(0,5),
     lwd=2, main="Bulk Density Pedotransfer Functions")
lines(density(BD_test$Saini),col="green", lwd=2)
lines(density(BD_test$Drew),col="red", lwd=2)
lines(density(BD_test$Jeffrey),col="cyan", lwd=2)
lines(density(BD_test$Grigal),col="orange", lwd=2)
lines(density(BD_test$Adams),col="magenta", lwd=2)
lines(density(BD_test$Honeyset_Ratkowsky),col="blue", lwd=2)
legend("topleft",
       legend = c("Original", "Saini", "Drew", "Jeffrey", "Grigal", "Adams",
                  "Honeyset_Ratkowsky"),
       fill=c("black", "green", "red", "cyan", "orange","magenta", "blue"))

# Plot the Selected function again
plot(density(BD_test$BD_test),type="l",col="black", ylim=c(0,3.5),
     lwd=2, main="Bulk Density Selected Function")
lines(density(BD_test$Honeyset_Ratkowsky),col="blue", lwd=2)
legend("topleft",legend = c("Original", "Honeyset_Ratkowsky"),
       fill=c("black", "blue"))

```

The PTF to be chosen for estimating the BD of the missing horizons should be the closest to the measured BD values. Once, the appropriate PTF was chosen, the `estimateBD` function is applied in the dataframe `dat` that was created at the end of the quality check. Here, new bd values are estimated for the rows in which the column 'bd' has missing values. Finally, a plot is generated to visualize the gap-filled bulk density values.

```{r, eval = T}
## 5.4 Estimate BD for the missing horizons ------------------------------------
dat$bd[is.na(dat$bd)] <-
  estimateBD(dat[is.na(dat$bd),]$soc, method="Honeyset_Ratkowsky1989")

# Explore the results
summary(dat$bd)
plot(density(BD_test$BD_test),type="l",col="black", ylim=c(0,3.5),
     lwd=2, main="Bulk Density Gap-Filling")
lines(density(dat$bd, na.rm = TRUE), col="green", lwd=2)
legend("topleft",legend = c("Original", "Original+Estimated"),
       fill=c("black", "green"))

```

## Check for outliers

Unrealistically high or low values can have considerable impact on the statistical analysis and thus it is key to identify and carefully check those values in order to get valid results and eliminate potential bias. Again, the summary() function is apt to show general descriptive statistics such as maxima or minima. Based on this assessment, more detailed views of the suspicious values can be obtained by filtering values above or below a certain threshold as done in the code below for soil organic carbon (SOC) values above 10 percent. If such values don't belong to soil types that would justify such exceptionally high SOC values, e.g. organic soils (Histosols), these rows can be removed based on the profile ID. The same process should be repeated for all soil properties.
Such evaluation can also be conducted visually for several properties at the same time using the `tidyverse` and `ggplot` package that allows to plot boxplots for several soil properties at the same time. To get more information on tidyverse, please follow this link: https://r4ds.had.co.nz/. For a comprehensive overview of the functionalities of ggplot, a more sophisticated way of plotting, this book provides a good overview: http://www.cookbook-r.com/Graphs/.

```{r, eval = T}
## 5.5 - Explore outliers ------------------------------------------------------
# Outliers should be carefully explored and compared with literature values.
# Only if it is clear that outliers represent impossible or highly unlikely 
# values, they should be removed as errors.
# 
# Carbon content higher than 15% is only typical for organic soil (histosols)
# We will remove all atypically high SOC as outliers
summary(dat$soc)
outliers <- na.omit(dat$ProfID[dat$soc > 10])
head(outliers)
dat <- dat[dat$ProfID != 6915,]
dat <- dat[dat$ProfID != 7726,]

# Explore bulk density data, identify outliers
# remove layers with Bulk Density < 1 g/cm^3
low_bd_profiles <- na.omit(dat$ProfID[dat$bd<1])
dat <- dat[!(dat$ProfID %in% low_bd_profiles),]

# Explore data, identify outliers
x <- pivot_longer(dat, cols = ph:cec, values_to = "value",
                  names_to = "soil_property")
x <- na.omit(x)
ggplot(x, aes(x = soil_property, y = value, fill = soil_property)) +
  geom_boxplot() + 
  facet_wrap(~soil_property, scales = "free")

```

## Harmonise soil layer depths
The last step towards a soil data table that can be used for mapping, is to harmonize the soil depth layers to 0-30 cm (or 30-60, or 60-100 cm respectively). This is necessary since we want to produce maps that cover exactly those depths and do not differ across soil profile locations. Thus, the relevant columns are selected from the dataframe, target soil properties, and upper and lower limit of the harmonised soil layer are specified (in depths).

In the following a new dataframe 'd' is created in which the standard depth layers are stored and named. The code below shows a for loop that calculates the values for the standard depth for each target soil property automatically using the ea_spline function of the 'ithir' package.

```{r, eval = T}
# 6 - Harmonize soil layers ====================================================
## 6.1 - Set target soil properties and depths ---------------------------------
names(dat)
dat <- select(dat, ProfID, HorID, x, y, top, bottom, ph, k, soc, bd, cec)

target <- c("ph", "k", "soc",  "bd", "cec")
depths <- t(c(0,30))

## 6.2 - Create standard layers ------------------------------------------------
d <- unique(select(dat, ProfID, x, y))

for (i in seq_along(target)) {
  vlow <- min(dat[,target[i]][[1]], na.rm = TRUE)
  vhigh <- max(dat[,target[i]][[1]], na.rm = TRUE)
  o <- dat[,c("ProfID", "top", "bottom",target[i])] %>% 
    na.omit() %>%
    as.data.frame(stringsAsFactors = FALSE)
  x <- ithir::ea_spline(obj = o, var.name = target[i], d = depths, 
                        vlow = vlow[[1]], vhigh = vhigh[[1]])$harmonised 
  x[x==-9999] <- NA
  x <- x %>% 
    as_tibble() %>% 
    select(-`soil depth`)
  
  names(x) <- c("ProfID",paste0(target[i],c("_0_30","_30_60","_60_100")))
  d <- d %>% left_join(x, by = "ProfID" )
}
d

```

## Harmonise units
Units are of paramount importance to deliver a high-quality map product. Therefore, special attention needs to be paid to a correct conversion/harmonisation of units particularly if different spreadsheets are combined. The mandatory soil properties need to be delivered in the following units:

```{r units}
# 7 - Harmonise units ==========================================================
#Harmonise units if different from target units
# Mandatory Soil Properties and corresponding units:
# Total N - ppm
# Available P - ppm
# Available K - ppm
# Cation exchange capacity cmolc/kg
# pH
# SOC - %
# Bulk density g/cm3
# Soil fractions (clay, silt and sand) - 

```

In the following, the available Potassium measurements from the soil profile data is converted from cmol~c~/kg to ppm. In addition, total N is converted from percent to ppm and the soil texture class values from g/kg to percent.

```{r, eval = T}

# Units soil profile data (dataframe d)
# 
head(d) # pH; K cmolc/kg; SOC %; BD g/cm3; CEC  cmolc/kg

# K => convert cmolc/kg to ppm (K *10 * 39.096)
d$k_0_30 <- d$k_0_30*10 * 39.096

head(chem)# P ppm; N %; K ppm
# N => convert % to ppm (N * 10000)
chem$tn <-chem$tn*10000

head(phys)# clay, sand, silt g/kg
# convert g/kg to % (/10)
phys$clay_0_30 <-phys$clay_0_30/10
phys$sand_0_30  <-phys$sand_0_30 /10
phys$silt_0_30 <-phys$silt_0_30/10

```

Finally, the different spreadsheets are merged into one single dataframe. For that, it is important to have matching column names in the dataframes that are to be merged.

```{r, eval = T}
# Add chemical and physical properties from additional datasets ==========================
  

# Rename columns to match the main data set
names(d)
names(chem)[1] <- 'ProfID'
names(chem)[4] <- 'p_0_30'
names(chem)[5] <- 'k_0_30' 
names(chem)[6] <- 'n_0_30'


#The chem dataframe comes from and independent dataset we need to create new unique ProfIDs 
#Create unique ProfID 
chem$ProfID <- seq(max(d$ProfID)+1,max(d$ProfID)+1+nrow(chem)-1)

# Add the new data as new rows using data.table we can add empty rows
# automatically for the not measured properties in the chem dataset
d <- rbind(setDT(d),setDT(chem), fill = TRUE)

#The phys dataframe with the texture instead shares the same ProfIDs (we can directly merge)
d <- merge(d, phys, by=c('ProfID', 'x', 'y'), all.x =T)

```

## Save the results
Before finalising the soil data preparation, it is recommendable to check again visually if the calculations were conducted correctly. Again, the combination of tidyverse and ggplot functions provides high efficiency and versatility to visualise figures with the desired soil properties. At last, the write_csv() function is used to save the dataframe as a .csv file in the Outputs folder (02-Outputs). With this, the soil data preparation is finalised.

```{r, eval = F}
# 8 - Plot  and save results ===================================================
names(d)
x <- pivot_longer(d, cols = ph_0_30:silt_0_30, values_to = "value",
                  names_sep = "_", 
                  names_to = c("soil_property", "top", "bottom"))
x <- mutate(x, depth = paste(top, "-" , bottom))
#x <- na.omit(x)
ggplot(x, aes(x = depth, y = value, fill = soil_property)) +
  geom_boxplot() + 
  facet_wrap(~soil_property, scales = "free")

# save the data
write_csv(d, "02-Outputs/harmonized_soil_data.csv")
```

