# Step 3: Mapping continuous soil properties

## Setting up repeated k-fold cross validation

Cross validation is one of the most used methods in DSM for assessing the overall accuracy of the resulting maps (Step 8, Figure 3). Since this is implemented along with the model calibration step, we explain the process at this stage. 

Cross validation consists of randomly splitting the input data into a training set and a testing set. However, a unique testing dataset can bias the overall accuracy. Therefore, k-fold cross validation randomly splits the data into k parts, using 1/k part of it for testing and k-1/k part for training the model. In order to make the final model more robust in terms of parameter estimations, we include repetitions of this process. The final approach is called repeated k-fold cross-validation, where k will be equal to ten in this process. A graphical representation of the 10-fold cross validation is shown in Figure \ref{fig:cv}. Note that green balls represent the samples belonging to the testing set and yellow balls are samples of the training set. Each row is a splitting step of the 10-folds, while each block (repetitions) represent the repetition step.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{images/cv.pdf}
  \caption{Schematic representation of the repeated cross-validation process.}
  \label{fig:cv}
\end{figure}

Step 5 in Figure 3 represents the repeated cross-validation, but note that after each single splitting step (the rows in Figure 4) the training data go to model calibration, which will be explained in Step 6 (next Section), and the testing data will be used with the calibrated model to produce the residuals (Step 8, Section 2.2.8). 
Repeated cross validation has been nicely implemented in the caret R package [@Kuhn2022], along with several calibration methods.

```{r, eval = F}
#_______________________________________________________________________________
#
# Quantile Regression Forest
# Soil Property Mapping
#
# GSP-Secretariat
# Contact: Isabel.Luotto@fao.org
#          Marcos.Angelini@fao.org
#_______________________________________________________________________________

#Empty environment and cache 
rm(list = ls())
gc()

# Content of this script =======================================================
# 0 - Set working directory, soil attribute, and packages
# 1 - Merge soil data with environmental covariates 
# 2 - Covariate selection
# 3 - Model calibration
# 4 - Uncertainty assessment
# 5 - Prediction
# 6 - Export final maps
#_______________________________________________________________________________


# 0 - Set working directory, soil attribute, and packages ======================

# Working directory
setwd('C:/GIT/Digital-Soil-Mapping')

# Load Area of interest (shp)
AOI <- '01-Data/AOI_Arg.shp'

# Terget soil attribute
soilatt <- 'k'

# Function for Uncertainty Assessment
load(file = "03-Scripts/eval.RData")

#load packages
library(tidyverse)
library(data.table)
library(caret)
library(quantregForest)
library(terra)
library(sf)
library(doParallel)

```



```{r, eval = F}
# 1 - Merge soil data with environmental covariates ============================

## 1.1 - Load covariates -------------------------------------------------------
# files <- list.files(path= '01-Data/covs/', pattern = '.tif$', full.names = T)
# covs <- rast(files)
# ncovs <- str_remove(files, "01-Data/covs/")
# ncovs <- str_remove(ncovs, ".tif")
# ncovs <- str_replace(ncovs, "-", "_")
# names(covs) <- ncovs 

covs <- rast("01-Data/covs/covs.tif")
ncovs <- names(covs)

## 1.2 - Load the soil data (Script 2) -----------------------------------------
dat <- read_csv("01-Data/data_with_coord.csv")

# Convert soil data into a spatial object (check https://epsg.io/6204)
dat <- vect(dat, geom=c("x", "y"), crs = crs(covs))

# Reproject point coordinates to match coordinate system of covariates
dat <- terra::project(dat, covs)
names(dat)

## 1.3 - Extract values from covariates to the soil points ---------------------
pv <- terra::extract(x = covs, y = dat, xy=F)
dat <- cbind(dat,pv)
dat <- as.data.frame(dat)

summary(dat)

## 1.4 - Target soil attribute + covariates ------------------------------------
d <- select(dat, soilatt, names(covs))
d <- na.omit(d)
```


```{r, eval = F}
# 2 - Covariate selection with RFE =============================================
## 2.1 - Setting parameters ----------------------------------------------------
# Repeatedcv = 3-times repeated 10-fold cross-validation
fitControl <- rfeControl(functions = rfFuncs,
                         method = "repeatedcv",
                         number = 10,         ## 10 -fold CV
                         repeats = 3,        ## repeated 3 times
                         verbose = TRUE,
                         saveDetails = TRUE, 
                         returnResamp = "all")

# Set the regression function
fm = as.formula(paste(soilatt," ~", paste0(ncovs,
                                             collapse = "+")))

# Calibrate the model using multiple cores
cl <- makeCluster(detectCores()-1)
registerDoParallel(cl)


## 2.2 - Calibrate a RFE model to select covariates ----------------------------
covsel <- rfe(fm,
              data = d,  
              sizes = seq(from=10, to=length(ncovs)-1, by = 5),
              rfeControl = fitControl,
              verbose = TRUE,
              keep.inbag = T)
stopCluster(cl)

## 2.3 - Plot selection of covariates ------------------------------------------
trellis.par.set(caretTheme())
plot(covsel, type = c("g", "o"))

# Extract selection of covariates and subset covs
opt_covs <- predictors(covsel)

```

```{r, eval = F}
# 3 - QRF Model calibration ====================================================
## 3.1 - Update formula with the selected covariates ---------------------------
fm <- as.formula(paste(soilatt," ~", paste0(opt_covs, collapse = "+")))

# parallel processing
cl <- makeCluster(detectCores()-1)
registerDoParallel(cl)

## 3.2 - Set training parameters -----------------------------------------------
fitControl <- trainControl(method = "repeatedcv",
                           number = 10,         ## 10 -fold CV
                           repeats = 3,        ## repeated 3 times
                           savePredictions = TRUE)

# Tune mtry hyperparameters
mtry <- round(length(opt_covs)/3)
tuneGrid <-  expand.grid(mtry = c(mtry-5, mtry, mtry+5))

## 3.3 - Calibrate the QRF model -----------------------------------------------
model <- caret::train(fm,
                      data = d,
                      method = "qrf",
                      trControl = fitControl,
                      verbose = TRUE,
                      tuneGrid = tuneGrid,
                      keep.inbag = T,
                      importance = TRUE)
stopCluster(cl)
gc()


## 3.4 - Extract predictor importance as relative values (%)
x <- randomForest::importance(model$finalModel)
model$importance <- x
## 3.5 - Print and save model --------------------------------------------------
print(model)
saveRDS(model, file = paste0("02-Outputs/models/model_",soilatt,".rds"))
```


```{r, eval = F}
# 4 - Uncertainty assessment ===================================================
# extract observed and predicted values
o <- model$pred$obs
p <- model$pred$pred
df <- data.frame(o,p)

## 4.1 - Plot and save scatterplot --------------------------------------------- 
(g1 <- ggplot(df, aes(x = o, y = p)) + 
  geom_point(alpha = 0.1) + 
   geom_abline(slope = 1, intercept = 0, color = "red")+
  ylim(c(min(o), max(o))) + theme(aspect.ratio=1)+ 
  labs(title = soilatt) + 
  xlab("Observed") + ylab("Predicted"))
ggsave(g1, filename = paste0("02-Outputs/residuals_",soilatt,".png"), scale = 1, 
       units = "cm", width = 12, height = 12)

## 4.2 - Print accuracy coeficients --------------------------------------------
# https://github.com/AlexandreWadoux/MapQualityEvaluation
eval(o,p)

## 4.3 - Plot Covariate importance ---------------------------------------------
(g2 <- varImpPlot(model$finalModel, main = soilatt, type = 1))

png(filename = paste0("02-Outputs/importance_",soilatt,".png"), 
    width = 15, height = 15, units = "cm", res = 600)
g2
dev.off()
```

```{r, eval = F}
# 5 - Prediction ===============================================================
# Generation of maps (prediction of soil attributes) 
## 5.1 - Produce tiles ---------------------------------------------------------
r <-covs[[1]]
t <- rast(nrows = 5, ncols = 5, extent = ext(r), crs = crs(r))
tile <- makeTiles(r, t,overwrite=TRUE,filename="02-Outputs/tiles/tiles.tif")

## 5.2 - Predict soil attributes per tiles -------------------------------------
# loop to predict on each tile

for (j in seq_along(tile)) {
  gc()
  t <- rast(tile[j])
  covst <- crop(covs, t)
  
  
  # plot(r)# 
  pred_mean <- terra::predict(covst, model = model$finalModel, na.rm=TRUE,  
                              cpkgs="quantregForest", what=mean)
  pred_sd <- terra::predict(covst, model = model$finalModel, na.rm=TRUE,  
                            cpkgs="quantregForest", what=sd)  
  
  
  
  # ###### Raster package solution (in case terra results in many NA pixels)
  # library(raster)
  # covst <- stack(covst)
  # class(final_mod$finalModel) <-"quantregForest"
  # # Estimate model uncertainty
  # pred_sd <- predict(covst,model=final_mod$finalModel,type=sd)
  # # OCSKGMlog prediction based in all available data
  # pred_mean <- predict(covst,model=final_mod)
  # 
  # 
  # ##################################  
  
  writeRaster(pred_mean, 
              filename = paste0("02-Outputs/tiles/soilatt_tiles/",
                                soilatt,"_tile_", j, ".tif"), 
              overwrite = TRUE)
  writeRaster(pred_sd, 
              filename = paste0("02-Outputs/tiles/soilatt_tiles/",
                                soilatt,"_tileSD_", j, ".tif"), 
              overwrite = TRUE)
  
  rm(pred_mean)
  rm(pred_sd)
  
  
  print(paste("tile",tile[j]))
}

## 5.3 - Merge tiles both prediction and st.Dev --------------------------------
f_mean <- list.files(path = "02-Outputs/tiles/soilatt_tiles/", 
                     pattern = paste0(soilatt,"_tile_"), full.names = TRUE)
f_sd <- list.files(path = "02-Outputs/tiles/soilatt_tiles/", 
                   pattern =  paste0(soilatt,"_tileSD_"), full.names = TRUE)
r_mean_l <- list()
r_sd_l <- list()

for (g in 1:length(f_mean)){
  r <- rast(f_mean[g])
  r_mean_l[g] <-r
  rm(r)
}

for (g in 1:length(f_sd)){
  
  r <- rast(f_sd[g])
  r_sd_l[g] <-r
  rm(r)
}
r_mean <-sprc(r_mean_l)
r_sd <-sprc(r_sd_l)
pred_mean <- mosaic(r_mean)
pred_sd <- mosaic(r_sd)

AOI <- vect(AOI)
pred_mean <- mask(pred_mean,AOI)
pred_sd <- mask(pred_sd,AOI)


plot(pred_mean)
plot(pred_sd)


```


```{r, eval = F}
# 6 - Export final maps ========================================================
## 6.1 - Mask croplands --------------------------------------------------------
msk <- rast("01-Data/mask_arg.tif")
plot(msk)
pred <- mask(pred_mean, msk)
plot(pred)

## 6.2 - Save results ----------------------------------------------------------
writeRaster(pred_mean, 
            paste0("02-Outputs/maps/",soilatt,"_QRF.tif"),
            overwrite=TRUE)
writeRaster(pred_sd, 
            paste0("02-Outputs/maps/",soilatt,"_QRF_SD.tif"),
            overwrite=TRUE)
```


## Model calibration

The model calibration step involves the use of a statistical model to find the relations between soil observations and environmental covariates. One of the most widely used models in DSM is random forest [@Breiman2001]. Random forest is considered a machine learning method which belongs to the decision-tree type of model. Random forest creates an ensemble of trees using a random selection of covariate. The prediction of a single tree is made based on the observed samples mean in the leaf. The random forest prediction is made by taking the average of the predictions of the single trees. The size of the number of covariates at each tree (mtry) can be fine-tuned before calibrating the model. 

Quantile regression forests (QRF, @Meinshausen2006) are a generalisation of the random forest models, capable of not only predicting the conditional mean, but also the conditional probability density function. This feature allows one to estimate the standard deviation of the prediction, as well as the likelihood of the target variable falling below a given threshold. In a context where a minimum level of a soil nutrient concentration may be decisive for improving the crop yield, this feature can play an important role for the GSNmap initiative. 

Model calibration will be implemented using the caret package [@Kuhn2022]. While we suggest to use QRF, caret provides a large set of models https://topepo.github.io/caret/available- models.html#) that might perform better in specific cases. In this regard, it is up to the user to implement a different model, ensuring  the product specifications (Section Product Specifications).

## Predicting soil attributes

After calibrating the model, caret will select the best set of parameters and will fit the model using the whole dataset. Then, the final model can be used to predict the target soil properties. The process uses the model and the values of the covariates at target locations. This is generally done by using the same input covariates as a multilayer raster format, ensuring that the names of the layers are the same as the covariates in the calibration dataset. In this step we will predict the conditional mean and conditional standard deviation at each raster cell.