---
title: "GSNmap report"
subtitle: "Country name"
aothor: "Author1, Author2, et al."
output: pdf_document
date: "2022-10-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 
<!-- 300 words max -->

Contents of the summary:
*   data source(s) (i.e. authors, year) 
*   Methodology 



# Soil Attribute: Potassium
## Descriptive statistics of input data

```{r load data, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(terra)
library(caret)
library(knitr)
library(summarytools)
# setwd("C:/GIT/GSNmap-TM/Digital-Soil-Mapping/")
load(file = "03-Scripts/eval.RData")

soilatt <- "k_0_30"

dxy <- read_csv("02-Outputs/harmonized_soil_data.csv")
model <- read_rds(paste0("02-Outputs/models/model_",soilatt,".rds"))
```

```{r input data, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
st_options(plain.ascii = FALSE, style = "rmarkdown")
options(width = 60)
kable(dxy[1:20,], digits = 3)
dfSummary(dxy[,soilatt], 
          plain.ascii  = FALSE,
          style        = 'grid',
          graph.magnif = 0.85,
          varnumbers = FALSE,
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")
```

## Resulting maps

### Predicted mean

```{r pred_mean, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
pred_mean <- rast(paste0("02-Outputs/maps/",soilatt,"_QRF.tif"))
terra:::plot(pred_mean, axes = TRUE, main = paste("Predicted mean", soilatt))
```

### Predicted standard deviation

```{r pred_sd, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
pred_sd <- rast(paste0("02-Outputs/maps/",soilatt,"_QRF_SD.tif"))
terra:::plot(pred_sd, axes = TRUE, 
             main = paste("Predicted standard deviation", soilatt))

```

## Accuracy assessment

### Observed vs Predicted

```{r scatterplot, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
o <- model$pred$obs
p <- model$pred$pred
df <- data.frame(o,p)

## 4.1 - Plot and save scatterplot --------------------------------------------- 
ggplot(df, aes(x = o, y = p)) + 
  geom_point(alpha = 0.1) + 
   geom_abline(slope = 1, intercept = 0, color = "red")+
  ylim(c(min(o), max(o))) + theme(aspect.ratio=1)+ 
  labs(title = soilatt) + 
  xlab("Observed") + ylab("Predicted")
```

### Accuracy indicators

```{r accuracy, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
kable(eval(p,o)[,1:6])

```

### Importance of covariates

```{r imp, echo=FALSE, fig.height=9, fig.width=6, message=FALSE, warning=FALSE, results='asis'}
options(width = 60)
# layout(matrix(1), widths = lcm(5), heights = lcm(10))
randomForest::varImpPlot(model$finalModel, main = soilatt, type = 1, )
```


<!-- --------------------------- NEW SOIL ATTRIBUTE ------------------------ -->
<!-- ----------------------------------------------------------------------- -->
# Soil Attribute: Available Phosphorous (Bray)
## Descriptive statistics of input data








