---
title: "GSNmap report"
subtitle: "Country name"
aothor: "Author1, Author2, et al."
output: word_document
date: "2022-10-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 
<!-- 300 words max -->

Contents of the summary:

* Specify the data products generated (mandatory & additional)

All national maps were generated following the Country guidelines and technical specifications for
Global Soil Nutrient and Nutrient Budget Maps (GSNmap) Phase I (FAO, 2022). 
All layers and accompanying standard deviation layers were created using a digital soil mapping approach based on the Quantile Regression Forest algorithm. 

The following table (Table 1) summarizes the data points used for generating a national GSNmap product.
```{r, echo =FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(terra)
library(caret)
library(knitr)

#Load data
dxy <- read.csv("../02-Outputs/harmonized_soil_data.csv")
load(file = "../03-Scripts/eval.RData")

table1 <- read.csv('National Report Table 1.csv')

#Data info table to edit directly in the word document
kable(table1, col.names = c("Data type" ,  "Source", "Number of points" ,"Year(s)"))

```




# Soil Attribute: Total Nitrogen
## Descriptive statistics of input data

```{r  message=FALSE, warning=FALSE, include=FALSE}


# target_properties<- c("ph_0_30", "k_0_30" , "soc_0_30" ,"bd_0_30", "cec_0_30","p_0_30",   
#                       "n_0_30","clay_0_30", "sand_0_30" ,"silt_0_30")

soilatt <- "n_0_30"

```

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}




model <- read_rds(paste0("../02-Outputs/models/model_",soilatt,".rds"))
  
  dat <-as.data.frame(dxy[,soilatt])
  names(dat)<-soilatt
  
  
if (soilatt == 'ph_0_30'){
  unit <-'NA'
  name <- 'pH'
}else if (soilatt == 'k_0_30'){
  unit <-'ppm'
  name <- 'Available Potassium'
}else if (soilatt == 'soc_0_30'){
  unit <-'percent'
  name <- 'Soil Organic Carbon'
}else if (soilatt == 'clay_0_30'){
  unit <-'percent'
   name <- 'Clay'
}else if (soilatt == 'bd_0_30'){
  unit <-'g/cm3'
  name <- 'Bulk Density'
}else if (soilatt == 'cec_0_30'){
  unit <-'cmolc/kg'
   name <- 'Cation Exchange Capacity'
}else if (soilatt == 'p_0_30'){
  name <- 'Available Phosphorus'
  unit <-'ppm'
}else if (soilatt == 'n_0_30'){
  name <- 'Total Nitrogen'
  unit <-'ppm'
}else if (soilatt == 'sand_0_30'){
  name <- 'Sand'
  unit <-'percent'
}else if (soilatt == 'silt_0_30'){
  name <- 'Silt'
  unit <-'percent'
}

  

 
  
  table2 <- data.frame(
  mean = round(mean(dat[,1],na.rm =T),1),
  min =  round(min(dat[,1],na.rm=T),1),
  median = round(median(dat[,1],na.rm=T),1),
  max =  round(max(dat[,1],na.rm=T),1),
  npoint = length(dat[complete.cases(dat),]),
  unit=unit
  )
  
kable(table2,col.names =c("mean",   "min" ,   "median" ,"max"  ,  "Number of points", "Unit"  ),
      caption = paste('Overview of the input data used for the', name, 'soil property map'),)  
  
  


```

## Resulting maps

### Predicted mean

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
pred_mean <- rast(paste0("../02-Outputs/maps/ISO_GSNmap_Ntot_Map030.tiff"))
terra:::plot(pred_mean, axes = TRUE, main = paste("Predicted mean", name))
```

### Predicted standard deviation

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
pred_sd <- rast(paste0("../02-Outputs/maps/ISO_SD_GSNmap_Ntot_Map030.tiff"))
terra:::plot(pred_sd, axes = TRUE, 
             main = paste("Predicted standard deviation", name))

```

## Accuracy assessment

### Observed vs Predicted

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
o <- model$pred$obs
p <- model$pred$pred
df <- data.frame(o,p)



## 4.1 - Plot and save scatterplot --------------------------------------------- 
ggplot(df, aes(x = o, y = p)) + 
  geom_point(alpha = 0.1) + 
   geom_abline(slope = 1, intercept = 0, color = "red")+
  ylim(c(min(o), max(o))) + theme(aspect.ratio=1)+ 
  labs(title = name) + 
  xlab("Observed") + ylab("Predicted")
```

### Accuracy indicators

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
kable(eval(p,o)[,1:6], caption = paste0('Accuracy Indicators: ',name))

```

### Importance of the covariates

```{r  echo=FALSE, fig.height=9, fig.width=6, message=FALSE, warning=FALSE, results='asis'}
options(width = 60)
# layout(matrix(1), widths = lcm(5), heights = lcm(10))
randomForest::varImpPlot(model$finalModel, main = paste0('Importance of the covariates: ',name), type = 1, )
```



<!-- --------------------------- NEW SOIL ATTRIBUTE ------------------------ -->
<!-- ----------------------------------------------------------------------- -->
# Soil Attribute: Available Phosphorous (Bray)
## Descriptive statistics of input data


```{r  message=FALSE, warning=FALSE, include=FALSE}


# target_properties<- c("ph_0_30", "k_0_30" , "soc_0_30" ,"bd_0_30", "cec_0_30","p_0_30",   
#                       "n_0_30","clay_0_30", "sand_0_30" ,"silt_0_30")

soilatt <- "p_0_30"

```

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}




model <- read_rds(paste0("../02-Outputs/models/model_",soilatt,".rds"))
  
  dat <-as.data.frame(dxy[,soilatt])
  names(dat)<-soilatt
  
  
if (soilatt == 'ph_0_30'){
  unit <-'NA'
  name <- 'pH'
}else if (soilatt == 'k_0_30'){
  unit <-'ppm'
  name <- 'Available Potassium'
}else if (soilatt == 'soc_0_30'){
  unit <-'percent'
  name <- 'Soil Organic Carbon'
}else if (soilatt == 'clay_0_30'){
  unit <-'percent'
   name <- 'Clay'
}else if (soilatt == 'bd_0_30'){
  unit <-'g/cm3'
  name <- 'Bulk Density'
}else if (soilatt == 'cec_0_30'){
  unit <-'cmolc/kg'
   name <- 'Cation Exchange Capacity'
}else if (soilatt == 'p_0_30'){
  name <- 'Available Phosphorus'
  unit <-'ppm'
}else if (soilatt == 'n_0_30'){
  name <- 'Total Nitrogen'
  unit <-'ppm'
}else if (soilatt == 'sand_0_30'){
  name <- 'Sand'
  unit <-'percent'
}else if (soilatt == 'silt_0_30'){
  name <- 'Silt'
  unit <-'percent'
}

  

 
  
  table2 <- data.frame(
  mean = round(mean(dat[,1],na.rm =T),1),
  min =  round(min(dat[,1],na.rm=T),1),
  median = round(median(dat[,1],na.rm=T),1),
  max =  round(max(dat[,1],na.rm=T),1),
  npoint = length(dat[complete.cases(dat),]),
  unit=unit
  )
  
kable(table2,col.names =c("mean",   "min" ,   "median" ,"max"  ,  "Number of points", "Unit"  ),
      caption = paste('Overview of the input data used for the', name, 'soil property map'),)  
  
  


```

## Resulting maps

### Predicted mean

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
pred_mean <- rast(paste0("../02-Outputs/maps/ISO_GSNmap_Pav_Map030.tiff"))
terra:::plot(pred_mean, axes = TRUE, main = paste("Predicted mean", name))
```

### Predicted standard deviation

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
pred_sd <- rast(paste0("../02-Outputs/maps/ISO_SD_GSNmap_Pav_Map030.tiff"))
terra:::plot(pred_sd, axes = TRUE, 
             main = paste("Predicted standard deviation", name))

```

## Accuracy assessment

### Observed vs Predicted

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
o <- model$pred$obs
p <- model$pred$pred
df <- data.frame(o,p)



## 4.1 - Plot and save scatterplot --------------------------------------------- 
ggplot(df, aes(x = o, y = p)) + 
  geom_point(alpha = 0.1) + 
   geom_abline(slope = 1, intercept = 0, color = "red")+
  ylim(c(min(o), max(o))) + theme(aspect.ratio=1)+ 
  labs(title = name) + 
  xlab("Observed") + ylab("Predicted")
```

### Accuracy indicators

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
kable(eval(p,o)[,1:6], caption = paste0('Accuracy Indicators: ',name))

```

### Importance of the covariates

```{r  echo=FALSE, fig.height=9, fig.width=6, message=FALSE, warning=FALSE, results='asis'}
options(width = 60)
# layout(matrix(1), widths = lcm(5), heights = lcm(10))
randomForest::varImpPlot(model$finalModel, main = paste0('Importance of the covariates: ',name), type = 1, )
```



<!-- --------------------------- NEW SOIL ATTRIBUTE ------------------------ -->
<!-- ----------------------------------------------------------------------- -->
# Soil Attribute: Soil Organic Carbon 
## Descriptive statistics of input data


```{r  message=FALSE, warning=FALSE, include=FALSE}


# target_properties<- c("ph_0_30", "k_0_30" , "soc_0_30" ,"bd_0_30", "cec_0_30","p_0_30",   
#                       "n_0_30","clay_0_30", "sand_0_30" ,"silt_0_30")

soilatt <- "soc_0_30"

```

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}




model <- read_rds(paste0("../02-Outputs/models/model_",soilatt,".rds"))
  
  dat <-as.data.frame(dxy[,soilatt])
  names(dat)<-soilatt
  
  
if (soilatt == 'ph_0_30'){
  unit <-'NA'
  name <- 'pH'
}else if (soilatt == 'k_0_30'){
  unit <-'ppm'
  name <- 'Available Potassium'
}else if (soilatt == 'soc_0_30'){
  unit <-'percent'
  name <- 'Soil Organic Carbon'
}else if (soilatt == 'clay_0_30'){
  unit <-'percent'
   name <- 'Clay'
}else if (soilatt == 'bd_0_30'){
  unit <-'g/cm3'
  name <- 'Bulk Density'
}else if (soilatt == 'cec_0_30'){
  unit <-'cmolc/kg'
   name <- 'Cation Exchange Capacity'
}else if (soilatt == 'p_0_30'){
  name <- 'Available Phosphorus'
  unit <-'ppm'
}else if (soilatt == 'n_0_30'){
  name <- 'Total Nitrogen'
  unit <-'ppm'
}else if (soilatt == 'sand_0_30'){
  name <- 'Sand'
  unit <-'percent'
}else if (soilatt == 'silt_0_30'){
  name <- 'Silt'
  unit <-'percent'
}

  

 
  
  table2 <- data.frame(
  mean = round(mean(dat[,1],na.rm =T),1),
  min =  round(min(dat[,1],na.rm=T),1),
  median = round(median(dat[,1],na.rm=T),1),
  max =  round(max(dat[,1],na.rm=T),1),
  npoint = length(dat[complete.cases(dat),]),
  unit=unit
  )
  
kable(table2,col.names =c("mean",   "min" ,   "median" ,"max"  ,  "Number of points", "Unit"  ),
      caption = paste('Overview of the input data used for the', name, 'soil property map'),)  
  
  


```

## Resulting maps

### Predicted mean

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
pred_mean <- rast(paste0("../02-Outputs/maps/ISO_GSNmap_SOC_Map030.tiff"))
terra:::plot(pred_mean, axes = TRUE, main = paste("Predicted mean", name))
```

### Predicted standard deviation

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
pred_sd <- rast(paste0("../02-Outputs/maps/ISO_SD_GSNmap_SOC_Map030.tiff"))
terra:::plot(pred_sd, axes = TRUE, 
             main = paste("Predicted standard deviation", name))

```

## Accuracy assessment

### Observed vs Predicted

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
o <- model$pred$obs
p <- model$pred$pred
df <- data.frame(o,p)



## 4.1 - Plot and save scatterplot --------------------------------------------- 
ggplot(df, aes(x = o, y = p)) + 
  geom_point(alpha = 0.1) + 
   geom_abline(slope = 1, intercept = 0, color = "red")+
  ylim(c(min(o), max(o))) + theme(aspect.ratio=1)+ 
  labs(title = name) + 
  xlab("Observed") + ylab("Predicted")
```

### Accuracy indicators

```{r  echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
kable(eval(p,o)[,1:6], caption = paste0('Accuracy Indicators: ',name))

```

### Importance of the covariates

```{r  echo=FALSE, fig.height=9, fig.width=6, message=FALSE, warning=FALSE, results='asis'}
options(width = 60)
# layout(matrix(1), widths = lcm(5), heights = lcm(10))
randomForest::varImpPlot(model$finalModel, main = paste0('Importance of the covariates: ',name), type = 1, )
```



# References

*FAO.* 2022. Country guidelines and technical specifications for global soil nutrient and nutrient budget maps GSNmap: Phase 1. Rome. https://doi.org/10.4060/cc1717en 





